---
title: "p8105_hw1"
author: "Qianying Wu"
date: "2023-09-20"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(moderndive)
library(ggplot2)
library(tidyverse)
```

## Problem 1

The dataset early_january_weather keeps track of the weather conditions in early January.  
```{r}
data("early_january_weather")
head(early_january_weather)
nrow(early_january_weather)
ncol(early_january_weather)

# Mean Temp
mean(early_january_weather$temp, na.rm = TRUE)
```
### Variables:
There are 15 variables in the data.
* Origin: Contains a 3-letter character.
* Year: integer variable representing year in the date.
* Month: integer variable representing month in the date.
* Day: integer variable representing day in the date.
* Hour: integer variable representing hour in one day. 
* Temp: Numeric variable with two decimal points, representing the temperature.
* Dewp: Numeric variable with two decimal points. 
* Humid: Numeric variable with two decimal points.
* Wind_dir: Numeric variable with some NAs. 
* Wind_speed: Variable with numeric values, representing speed of wind.
* Wind_gust: Variable with numeric values and lots of NAs. 
* Precip: Variable with numeric values. Here all of them are 0.0.
* Pressure: Variable with numeric values and some NAs.
* Visib: Variable with numeric values.
* Time-hour: Variable with date and time.

### Size of the Data
There are 358 rows and 15 columns in the data.

### Mean Temperature
The mean temperature for early_january_weather is 39.58212.

### Scatterplot
```{r}
scatterplot <- ggplot(data = early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() + 
  labs(x = "Time Hour", y = "Temperature", color = "Humidity") +
  theme_minimal()

print(scatterplot)

ggsave("scatterplot.png", scatterplot)
```

In this scatterplot, the x axis is time hour, y axis is the temperature, and the data are separated by different levels of humidity in different colors of blue. From the scatterplot, we can observe the overall trend of temperature is increasing. Also,  the temperature has small patterns of rise and fall as the time_hour proceeds. From the graph, the maximum temperature is around 58, which occurs slightly after time_hour of Jan 14. The minimum temperature is around 23, which is at the beginning of January. 
We can also observe when time_hour before Jan 07, the humidity tends to be low, and humidity becomes high in the time_hour of Jan 14. After Jan 14, the temperature begins to drop and humidity becomes low again. 

## Problem 2
```{r}
p2 <- tibble(num = rnorm(10),
             log_vec = num > 0,
             char = sample(letters, 10),
             factor = factor(sample(1:3, 10, replace = TRUE))
             )

mean(p2$num)
mean(p2$log_vec)
mean(p2$char)
mean(p2$factor)

# Using the pull function
p2 %>% pull(num) %>% mean()
p2 %>% pull(log_vec) %>% mean()
p2 %>% pull(char) %>% mean()
p2 %>% pull(factor) %>% mean()
```

* The numeric and logical vectors work when calculating their mean, while the character and factor vectors don't work.

```{r, results = "hide"}
as.numeric(p2$log_vec)
as.numeric(p2$char)
as.numeric(p2$factor)
```

From the result, we can see R can successfully apply `as.numeric` to logical and factor vectors.
When applying `as.numeric` to a logical variable, R will convert `TRUE` to 1 and `FALSE` to 0. 
When applying `as.numeric` to a factor variable, R will convert the level number to its underlying numbers. 
When applying `as.numeric` to a character variable, R failed to convert the characters into numbers, so there will be NAs generated.
Therefore, this explains why we can't get the mean of the character vector. For the factor vector, we can't get its mean because they represent levels rather than actual numbers. 

